---
title: "Task 4: Becoming a databender"
output:
  html_document:
    keep_md: true
    theme: cerulean
    code_folding: hide
    
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

library(nycflights13)
library(tidyverse)
```
## Background
This report was created in response to the following scenari;

"You just started your internship at a big firm in New York, and your manager gave you an extensive file of flights that departed JFK, LGA, or EWR in 2013. From this data (install.packages(nycflights13)) your manager wants you to answer the following questions;

  1. If I am leaving before noon, what are my top two airline options at each airport (JFK, LGA, EWR) that will have the least amount of delay time?
  
  2. Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?
  
  3. Which destination airport is the worst airport for arrival time?
"

### If I am leaving before noon, what are my top two airline options at each airport (JFK, LGA, EWR) that will have the least amount of delay time?<br><br>

In order to answer this question, I first filtered out of the data every flight that departed after 12:00 PM. Then for each of the origin airports, I created a new set of data that only included flights that left from that airport. I then used the "transmute" function to create new tables of data, one for each origin airport, that contained the columns "sched_dep_time", "carrier", and a new column called "total_delay" that contained the total delay time for each flight. All that was left to do was arrange the data so that the carriers that had the lowest average total delay time would appear at the top row. The lists found below are the result after arranging the data. 

The airlines with the lowest average delay time are Delta and American at JFK, US and American at LGA, and Alaskan and Virgin America at EWR. 

```{r, comment=NA, warning=FALSE}
early_jfk <- filter(flights, sched_dep_time <= 1200 & origin == "JFK")
q1 <- transmute(early_jfk,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.1 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
v1 <- arrange(summ1.1, avg_delay)

early_lga <- filter(flights, sched_dep_time <= 1200 & origin == "LGA")
q1 <- transmute(early_lga,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.2 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
v2 <- arrange(summ1.2, avg_delay)

early_ewr <- filter(flights, sched_dep_time <= 1200 & origin == "EWR")
q1 <- transmute(early_ewr,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.3 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
v3 <- arrange(summ1.3, avg_delay)

sidebyside <- function(..., width=60){
  l <- list(...)
  p <- lapply(l, function(x){
        xx <- capture.output(print(x, width=width))
        xx <- gsub("\"", "", xx)
        format(xx, justify="left", width=width)
      }
  )
  p <- do.call(cbind, p)
  sapply(seq_len(nrow(p)), function(x)paste(p[x, ], collapse=""))
}
t1 <- "JFK"
t2 <- "LGA"
t3 <- "EWR"

sidebyside(t1, t2, t3, width = 30)
sidebyside(v1, v2, v3, width = 30)

```
<br><br>

 

### Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?

For this question, I first created a new table of only Delta's flight data. From there, I grouped the data by the origin of flight. Then I summarized the data by average arrival delay time, and arranged it so that the airport with the lowest average was shown at the top of the list. The result can be seen below.

The JFK airport has the lowest average arrival delay time for all Delta flights, which means it's your best option for minimizing your chance of a late arrival when using Delta.

```{r, comment=NA}
on_time <- filter(flights, carrier == "DL")
q2 <- group_by(on_time, origin)
summ2 <- summarise(q2, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ2, avg_delay)
```


### Which destination airport is the worst airport for arrival time?

This question was relatively easy to answer. All I had to do was group together all of the destinations of all flights and summarize the data by average arrival time. Then I arranged the data in descending order so that the destination with the highest average arrival delay time was placed at the top. 

The Columbia Metropolitan Airport has the highest average arrival delay time out of all the destinations in the data. 

```{r, comment=NA}
q3 <- group_by(flights, dest)
summ3 <- summarise(q3, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ3, desc(avg_delay))
```


<!-- 
### If I am leaving before noon, what are my top two airline options at each airport (JFK, LGA, EWR) that will have the least amount of delay time?<br><br>

##### John F. Kennedy International Airport (JFK)
```{r, comment=NA}
early_jfk <- filter(flights, sched_dep_time <= 1200 & origin == "JFK")
q1 <- transmute(early_jfk,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.1 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.1, avg_delay)
```
<br><br>

##### LaGuardia Airport (LGA)
```{r, comment=NA}
early_lga <- filter(flights, sched_dep_time <= 1200 & origin == "LGA")
q1 <- transmute(early_lga,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.2 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.2, avg_delay)
```
<br><br>

##### Newark Liberty International Airport (EWR)
```{r, comment=NA}
early_ewr <- filter(flights, sched_dep_time <= 1200 & origin == "EWR")
q1 <- transmute(early_ewr,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.3 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.3, avg_delay)
```
<br><br>




### Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?

```{r, comment=NA}
on_time <- filter(flights, carrier == "DL")
q2 <- group_by(on_time, origin)
summ2 <- summarise(q2, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ2, avg_delay)
```


### Which destination airport is the worst airport for arrival time?

```{r, comment=NA}
q3 <- group_by(flights, dest)
summ3 <- summarise(q3, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ3, desc(avg_delay))
```

-->