---
title: "Task 4: Becoming a databender"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    
---

```{r, include=FALSE}
library(dplyr)
library(mosaic)
library(ggplot2)

library(nycflights13)
library(tidyverse)
```



### If I am leaving before noon, what are my top two airline options at each airport (JFK, LGA, EWR) that will have the least amount of delay time?<br><br>

##### John F. Kennedy International Airport (JFK)
```{r, comment=NA}
early_jfk <- filter(flights, sched_dep_time <= 1200 & origin == "JFK")
q1 <- transmute(early_jfk,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.1 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.1, avg_delay)
```
<br><br>

##### LaGuardia Airport (LGA)
```{r, comment=NA}
early_lga <- filter(flights, sched_dep_time <= 1200 & origin == "LGA")
q1 <- transmute(early_lga,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.2 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.2, avg_delay)
```
<br><br>

##### Newark Liberty International Airport (EWR)
```{r, comment=NA}
early_ewr <- filter(flights, sched_dep_time <= 1200 & origin == "EWR")
q1 <- transmute(early_ewr,
                sched_dep_time,
                carrier,
                total_delay = arr_delay + dep_delay
                )
q1g <- group_by(q1, carrier)
summ1.3 <- summarise(q1g, avg_delay = mean(total_delay, na.rm = TRUE))
arrange(summ1.3, avg_delay)
```
<br><br>




### Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?

```{r, comment=NA}
on_time <- filter(flights, carrier == "DL")
q2 <- group_by(on_time, origin)
summ2 <- summarise(q2, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ2, avg_delay)
```


### Which destination airport is the worst airport for arrival time?

```{r, comment=NA}
q3 <- group_by(flights, dest)
summ3 <- summarise(q3, avg_delay = mean(arr_delay, na.rm = TRUE))
arrange(summ3, desc(avg_delay))
```

